### 2019-02-07 paper/Bibliography.tex
bibliography added

\begin{thebibliography}{9}

\bibitem{gmm} 
C. Stauffer and W. Grimson. Adaptive background mixture
models for real-time tracking. In Computer Vision and Pattern Recognition, 1999. IEEE Computer Society Conference on., volume 2, pages 2 vol. (xxiii+637+663), 1999.

\end{thebibliography}