### 2019-02-07 paper/Bibliography.tex
citation added

\begin{thebibliography}{9}

\bibitem{gmm} 
C. Stauffer and W. Grimson. Adaptive background mixture
models for real-time tracking. In Computer Vision and Pattern Recognition, 1999. IEEE Computer Society Conference on., volume 2, pages 2 vol. (xxiii+637+663), 1999.

\bibitem{vibe}
O. Barnich and M. Van Droogenbroeck. Vibe: A universal
background subtraction algorithm for video sequences. Im-
age Processing, IEEE Transactions on, 20(6):1709 –1724,
june 2011.

\end{thebibliography}