### 2019-02-07 paper/Bibliography.tex
Added section Experiments. Images to fix.

\begin{thebibliography}{9}

\bibitem{gmm} 
C. Stauffer and W. Grimson. Adaptive background mixture
models for real-time tracking. In Computer Vision and Pattern Recognition, 1999.
IEEE Computer Society Conference on., volume 2, pages 2 vol. (xxiii+637+663), 1999.

\bibitem{vibe}
O. Barnich and M. Van Droogenbroeck. Vibe: A universal
background subtraction algorithm for video sequences. Im-
age Processing, IEEE Transactions on, 20(6):1709 –1724,
june 2011.

\bibitem{pbas_and_scene_analysis_fpga}
T. Kryjak, M. Komorkiewicz and M. Gorgon.
Real-time Foreground Object Detection Combining
the PBAS Background Modelling Algorithm
and Feedback from Scene Analysis Module.
INTL Journal of Electronics and Telecommunications, 2014, VOL. 60, NO. 1, PP. 61–72

\end{thebibliography}